# Time to start exploring the data!

```{r include=FALSE, echo=FALSE, eval=FALSE}
## Suppress this chunk in the book
install.packages('ggplot2')
install.packages('magrittr')
install.packages('Hmisc')
```

```{r}
require('magrittr')
require('ggplot2')
require('Hmisc')
```

So far we've loaded the dataset, we've done some high level exploration of the structure, cleaned up the data and even created a simple linear model. In most projects, this will be the hardest part. Once you have your data loaded and cleaned, a common next step is to use R's powerful graphics packages to create some _exploratory graphics_ to help you visualize the dataset.

## R Graphics

R has a few graphics packages to choose from. Many useful plots come 'out of the box' in R.  For example, if we want to see a scatterplot matrix of all the data columns in the _airquality_ dataset, we could use the _pairs_ funtion.

```{r}
pairs(airquality)
```


While the basic package is nice, we will be using _ggplot2_ today. It is a powerful graphics package that is designed to make your plots adaptable, scalable, and easy to produce.  It is a very popular graphics package used by a large number of R users worldwide.

Many times, it's helpful to just do a scatterplot of two variables

```{r}
p <- ggplot(airquality, aes(Ozone, Solar.R)) + geom_point()
p
```

# Exploring further

