# Appendix

## More on Dealing with Non-numeric data

So far, we've primarily worked with numeric data.

Let's talk briefly about how R handles string or character data.

Here's a data frame with some letters.

```{r}
myname <- data.frame(letter=c("k","e","l","s","e","y"))
```

Let's look at the structure of the _myname_ object, and the _letter_ field of the object.

```{r}
myname %>% str()
myname$letter %>% str()
```

Hmm, it says `Factor w/ 5 levels "e","k","l","s",..: 2 1 3 4 1 5`. What does this mean?

By default, R stores string fields in data frames as _factors_.

Behind the scenes, R identifies the unique character strings in the field, substitutes each one with a unique integer, and remembers the original strings as _levels_.

```{r}
myname$letter %>% levels()
myname$letter %>% as.integer()
```

For many processes and analyses, this data structure will work fine.

In some situations, however, it becomes problematic.

Let's try to modify the _myname_ data frame, and make the letters spell out _kelseysm_ rather than _kelsey_.

We need to add letters `s` and  `m` to the next rows. We can do that with the `rbind` function.

```{r}
myname <- myname %>% rbind("s")
myname <- myname %>% rbind("m")
```

An Error! That's not what we want. This happened because "m" is not a registerd factor for `letter`, so R won't add it.

There are a couple of ways to override R's default behavior of storing text fields as columns.

If we're creating a data frame from scratch, we can set the _stringsAsFactors_ parameter equal to _FALSE_.

```{r}
myname <- data.frame(letter=c("k","e","l","s","e","y"), stringsAsFactors=FALSE)
```

Check out the structure _str_ and see what is different.

Let's try making the same modifications now.

```{r}
myname <- myname %>% rbind("s")
myname <- myname %>% rbind("m")
```

The character data type is much more flexible.

If we've already created (or read) a data frame, we can use the _as.character_ function on the text field itself.

```{r}
myname <- data.frame(letter=c("k","e","l","s","e","y"))
myname %>% str()
myname$letter <- myname$letter %>% as.character()
myname %>% str()
myname <- myname %>% rbind("s")
myname <- myname %>% rbind("m")
```

Looking out for factors is critical when typecasting numeric data that is mistakenly stored as text.

Here's another data frame.

```{r}
fives <- data.frame(v1=c("5","10","15","20"))
```

Let's look at the structure and levels:

```{r}
fives %>% str()
fives$v1 %>% levels()
```

We want these data points stored as a numeric type, but currently, they're stored as text and, specifically, as a factor.

What does it look like if we convert the factor to numeric?

```{r}
fives$v1 %>% as.numeric()
```

So problematic! Why did this happen?

Because the factor levels are what is converted to numbers, not the characters inside the strings. Let's first convert factors to characters and then convert to numbers.

```{r}
fives$v1 %>% as.character() %>% as.numeric()
```

That works as expected.

Are factors _ever_ useful? Yes.

Here's some pretend survey data.

```{r}
survey <- data.frame(gender=rep(c("Male", "Female"), c(5, 7)), response=rep(c("Excellent", "Good", "Fair", "Poor"), 3))
```

Let's change the levels of the gender field.

```{r}
survey$gender %>% levels()
levels(survey$gender) <- c("F","M")
```

What does it look like now?

We can also take advantage of our understanding of the factor data type to convert the response variable into a numeric field.

I want _excellent_ to map to 4, _good_ to map to 3, _fair_ to map to 2 and _poor_ to map to 1.

We'll use the _as.numeric_ function eventually, but first we need to see if the levels are in the order that we want them.

```{r}
survey$response %>% levels()
```

They're not. Let's change them, so that less favorable responses will appear before more favorable responses when alphabetized.

After we change the levels, we'll need to turn the response field into a character field, and then re-factor.

```{r}
levels(survey$response) <- c("D Excellent","B Fair","C Good","A Poor")
survey$response <- survey$response %>% as.character() %>% as.factor()
survey$response %>% levels()
levels(survey$response) <- c("Poor","Fair","Good","Excellent")
```

Much better. Now we can call the _as.numeric_ function.

```{r}
survey %>% mutate(as.numeric(response))
```

How useful!

### The takeaway

By default, R stores text fields as factors. And factors are strange.

At times it seems like they are working against us.

But if we understand how they work, we can make them work for us!

Now, back to numeric data.

## Webscraping and text analysis

* [Webscraping and text analysis](http://datascienceplus.com/goodreads-webscraping-and-text-analysis-with-r-part-1/)
* [R-Bloggers search for web scraping](https://www.r-bloggers.com/search/web%20scraping/)
* [A tutorial using Super Bowl Data](https://rpubs.com/Radcliffe/superbowl)
* [Scraping with rvest](https://www.r-bloggers.com/rvest-easy-web-scraping-with-r/)
* [Scraping with Selenium](https://www.r-bloggers.com/scraping-with-selenium/)

## Machine Learning in R

* [Deep learning](https://www.r-bloggers.com/a-simple-workflow-for-deep-learning/)
* [Machine learning in R](https://www.r-bloggers.com/machine-learning-with-r/)
* [R for Machine Learning notes](https://ocw.mit.edu/courses/sloan-school-of-management/15-097-prediction-machine-learning-and-statistics-spring-2012/lecture-notes/MIT15_097S12_lec02.pdf)
* [R Machine Learning Group on Reddit](https://www.reddit.com/r/MachineLearning/comments/575jz8/discussion_join_us_on_rlearnmachinelearning/#bottom-comments)

## Interactive R Graphics

* [Interactive visualizations with R](http://ouzor.github.io/blog/2014/11/21/interactive-visualizations.html)
    * _"A minireview of R packages ggvis, rCharts, plotly and googleVis for interactive visualizations"_
* [Shiny Tutorial](http://shiny.rstudio.com/tutorial/)
* [Plotly ggplot](https://plot.ly/ggplot2/)
* [Flex Dashboard](http://rmarkdown.rstudio.com/flexdashboard/)

## Maps

* [Leaflet for R](http://rstudio.github.io/leaflet/)
    * Intro to making interactive maps in R
* [Spatial Data with R](http://robinlovelace.net/r/2014/01/30/spatial-data-with-R-tutorial.html)
* [Making Maps with R](http://www.molecularecologist.com/2012/09/making-maps-with-r/)
* [R-Bloggers search for maps](https://www.r-bloggers.com/search/maps/)

## Handling more data

* [Handling large data sets in R](https://rstudio-pubs-static.s3.amazonaws.com/72295_692737b667614d369bd87cb0f51c9a4b.html)
* [A quick tutorial](http://davetang.org/muse/2013/09/03/handling-big-data-in-r/)
* [Best practices for storing and using data frames too large for memory](http://stackoverflow.com/questions/1875795/best-practices-for-storing-and-using-data-frames-too-large-for-memory)
* [11 tips](http://theodi.org/blog/fig-data-11-tips-how-handle-big-data-r-and-1-bad-pun)
* [5 strategies](https://www.r-bloggers.com/five-ways-to-handle-big-data-in-r/)

## General

* [Task Views](https://cran.r-project.org/web/views/)
* [Data Structures](http://adv-r.had.co.nz/Data-structures.html)
